---
title: "Worked Example 2 - Outbreak Modelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example 2 - Outbreak Modelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This worked example demonstrates several ways the package can be used to model the occurrence and risk of yellow fever outbreaks. An "outbreak" is defined as taking place whenever a case of yellow fever is reported, and lasts as long as the time interval between new cases remains less than 10 days (the sum of the incubation and infectious periods used by the package).

The reporting of yellow fever cases is governed by the probability of an infection which causes severe but non-fatal symptoms being reported (p_rep_severe) and the probability of a fatal infection being reported (p_rep_death). It is assumed that infections not leading to severe or fatal symptoms are not reported. The proportion of infections which cause severe symptoms is given by the quantity p_severe_inf, and the proportion of severe infections which lead to death is given by the quantity p_death_severe_inf. These have values of 0.12 and 0.39 respectively within the package, based on [Servadio et al](https://doi.org/10.1186/s12879-021-06535-4), but can be set to other values for manual calculations, as shown below.

As described in [Worked Example 1 - Single Model Run](https://mrc-ide.github.io/YellowFeverDynamics/articles/AWorkedExampleAModelRun.html), we obtain data on the number of new infectious individuals at each time point during a modelled period from the case_data_generate() function:

```{r}
library(YellowFeverDynamics)
input_data <- readRDS(file = paste(path.package("YellowFeverDynamics"), "/exdata/input_data_example.Rds", sep = ""))
vacc_data <- input_data$vacc_data[1, , ]
pop_data <- input_data$pop_data[1, , ]  
FOI_spillover <- 5.0e-8             # Force of infection due to spillover from sylvatic reservoir
R0 <- 0.0                           # Basic reproduction number for urban spread of infection
model_results <- case_data_generate(FOI_spillover, R0, vacc_data, pop_data, year0 = input_data$years_labels[1], mode_start = 1, n_reps = 1, year_end = 2001, year_data_begin = 1999, vaccine_efficacy = 1.0, start_SEIRV = NULL, dt = 1.0)
```

We process results using the get_outbreak_data() function to produce information on all outbreaks and reported cases during the modelled time period.

```{r}
case_data = model_results$C[1,] # Vector of data points showing number of new cases at each time point [TBC]
year_data = model_results$year # Matching vector indicating the year for each time point [TBC]
p_rep_severe = 0.1 # The probability of a case with severe but non-fatal symptoms being reported
p_rep_death = 0.2 # The probability of a fatal case being reported
outbreak_data <- get_outbreak_data(case_data, year_data, p_rep_severe, p_rep_death)
```

get_outbreak_data() produces a list containing the number of outbreaks (n_outbreaks), data on each individual outbreak (outbreak_list - number of reported cases, start and end days and years), a time series showing the number of reported cases and deaths at each time point (rep_pts), and a time series showing the number of reported cases and deaths in each year (rep_annual).

```{r}
outbreak_data$outbreak_list
```

```{r}
outbreak_data$rep_annual
```

As noted in [Worked Example 1](https://mrc-ide.github.io/YellowFeverDynamics/articles/AWorkedExampleAModelRun.html), we can estimate the outbreak risk (the probability of one or more outbreaks, i.e. one or more reported cases) in a given time period by running the model a large number of times and calculating the proportion of repetitions in which one or more cases are reported.

```{r}
p_severe_inf <- 0.12
p_death_severe_inf <- 0.39
n_reps <- 100 # Number of repetitions
model_results_multi <- case_data_generate(FOI_spillover, R0, vacc_data, pop_data, year0 = input_data$years_labels[1], mode_start = 1, n_reps, year_end = 2001, year_data_begin = 1999, vaccine_efficacy = 1.0, start_SEIRV = NULL, dt = 1.0)
pts <- model_results_multi$year == 1999 # Time period for which to calculate outbreak risk, here defined as a particular year
outbreak_occurence_values <- rep(0, n_reps) # Vector of n_reps 0/1 values indicating whether an outbreak occurs for that repetition
for(rep in 1:n_reps){
  C_sum <- floor(sum(model_results_multi$C[rep, pts])) # Total new infections in time period
  C_severe_sum <- rbinom(1, C_sum, p_severe_inf)   # Total new severe infections
  C_death_sum <- rbinom(1, C_severe_sum, p_death_severe_inf) # Total fatal infections
  C_death_sum_rep <- rbinom(1, C_death_sum, p_rep_death)     # Total reported fatal infections
  C_sum_rep <- C_death_sum_rep + rbinom(1, C_severe_sum-C_death_sum, p_rep_severe) # Total reported infections
  if(C_sum_rep > 0){outbreak_occurence_values[rep] <- 1}
}
outbreak_risk <- mean(outbreak_occurence_values)
outbreak_risk
```

We can compute the outbreak risk in a year automatically and more efficiently using the outbreak_risk_generate() function. This function runs the time before the period of interest n_sets times, then models the period of interest n_reps times for each set, giving n_sets sets of values of annual outbreak risk:

```{r}
outbreak_risk <- outbreak_risk_generate(FOI_spillover, R0, vacc_data, pop_data, year0 = input_data$years_labels[1], mode_start = 1, n_sets=1, n_reps = 100, year_end = 2005, year_data_begin = 1999, vaccine_efficacy = 1.0, start_SEIRV = NULL, dt = 1.0, p_rep_severe, p_rep_death)
outbreak_risk
```
