---
title: "Worked Example 5 - Parameter Estimation Using Markov Chain Monte Carlo Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example 5 - Parameter Estimation Using Markov Chain Monte Carlo Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This worked example demonstrates how to use the package to estimate parameters of the yellow fever model from observed epidemiological data within a Bayesian framework using Markov chain Monte Carlo sampling.

We load input data, environmental data and the observed data to be used for comparison in the formats used in previous worked examples (see [Guide 2 - Inputs](https://mrc-ide.github.io/YellowFeverDynamics/articles/BGuideBInputs.html) for details of the input format, [Guide 3 - Calculating Parameters From Environmental Data](https://mrc-ide.github.io/YellowFeverDynamics/articles/BGuideCParametersEnvironment.html) for details of how environmental data is formatted and used to generate parameter values, and [Worked Example 3 - Generating Epidemiological Data Sets](https://mrc-ide.github.io/YellowFeverDynamics/articles/AWorkedExampleCDataGenerate.html) for details of observed data formats):

```{r,eval=FALSE}
library(YellowFeverDynamics)
input_data <- readRDS(file = paste(path.package("YellowFeverDynamics"), "/exdata/input_data_example.Rds", sep = ""))
enviro_data <- read.csv(file = paste(path.package("YellowFeverDynamics"), "/exdata/enviro_data_example.csv", sep = ""), header = TRUE)
n_env_vars = ncol(enviro_data)-1
obs_sero_data <- read.csv(file = paste(path.package("YellowFeverDynamics"), "/exdata/obs_sero_data_example.csv", sep = ""), header = TRUE)
obs_case_data <- read.csv(file = paste(path.package("YellowFeverDynamics"), "/exdata/obs_case_data_example.csv", sep = ""), header = TRUE)
obs_outbreak_data <- NULL
enviro_coeffs_ini <- as.numeric(read.csv(file = paste(path.package("YellowFeverDynamics"), "/exdata/enviro_coeffs_example.csv", sep = ""), header = TRUE)) # Values of environmental coefficients to use as initial parameter values
```

We set additional inputs as follows. The parameters varied in the chain are the natural logarithm values of the actual parameters of interest; the inputs log_params_min, log_params_max and log_params_ini are set accordingly. The maximum, minimum and initial values of the varied parameters must be entered into log_params_min, log_params_max and log_params_ini in the following order:

-Values of spillover FOI or the coefficients used to calculate spillover FOI from environmental data
-Values of R0 or the coefficients used to calculate R0 from environmental data (if R0 is to be varied)
-Vaccine efficacy (if varied)
-Severe non-fatal case reporting probability (if varied)
-Fatal case reporting probability (if varied)

The maximum and minimum values are not all necessarily used; if prior_type is set to "flat", then they are hard limits for all parameters. If prior_type is set to "zero" or "norm", only the limits for p_rep_severe and p_rep_death are used.

```{r,eval=FALSE}
Niter <- 1 #Number of iterations to run
type <- "FOI+R0 enviro" # Type of parameter set used to calculate spillover FOI and R0 (here both calculated from environmental covariates)
log_params_min <- c(rep(-30, n_env_vars), rep(-5, n_env_vars), log(c(0.9,0.01, 0.01))) # Minimum log parameter values (here FOI coefficients, R0 coefficients, vaccine efficacy, severe non-fatal case reporting probability, fatal case reporting probability)
log_params_max <- c(rep(-10, n_env_vars), rep(0, n_env_vars), log(c(1.0,0.25, 0.25))) # Maximum log parameter values
log_params_ini <- c(log(enviro_coeffs_ini), log(c(0.95,0.1,0.2))) # Initial log parameter values
n_reps <- 5 # Number of repetitions for which to run model and obtain average outputs
mode_start <- 1 # Flag indicating how to set initial population immunity level in addition to vaccination
                                    # If mode_start = 0, only vaccinated individuals
                                    # If mode_start = 1, shift some non-vaccinated individuals into recovered to give herd immunity
prior_type <- "norm" # Type of prior likelihood calculation to use
dt <- 5.0 # Time increment in days
R0_fixed_values = vaccine_efficacy = p_rep_severe = p_rep_death = NULL #Fixed R0 values, vaccine efficacy and severe and fatal case reporting probabilities; here set to NULL because R0, vaccine efficacy and reporting probabilities are variable parameters being estimated
filename_prefix <- "Chain" # Prefix to be used for .csv output files; a new file is created every 10,000 iterations with incremental number suffixes (Chain1.csv, Chain2.csv, etc.)
```

We then run the MCMC() function. Output data recorded at each iteration is saved to an output file every 10 iterations.

```{r,eval=FALSE}
MCMC(log_params_ini, input_data, obs_sero_data, obs_case_data, obs_outbreak_data, filename_prefix, Niter, type,
     log_params_min,log_params_max, n_reps, mode_start, prior_type, dt, enviro_data, R0_fixed_values, vaccine_efficacy,
     p_rep_severe,p_rep_death)
```

The saved output of MCMC() takes the form of a data frame of values of the current likelihood (posterior_current), likelihood calculated from proposed parameter values (posterior_prop), current and proposed parameter values, a flag indicating whether the proposed parameter values were accepted (set to 1 if they were accepted, 0 if they were not), and the chain covariance.

```{r,eval=FALSE}
chain_example <- read.csv(file = paste(path.package("YellowFeverDynamics"), "/exdata/Chain_example.csv", sep = ""), header = TRUE)
head(chain_example, 1)
```
