---
title: "Guide 3 - Calculating Parameters From Environmental Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guide 3 - Calculating Parameters From Environmental Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how the package generates values of epidemiological parameters (the force of infection for spillover from the sylvatic reservoir and the basic reproduction number for human-to-human transmission) from environmental covariates (including population, vegetation coverage, temperature suitability for yellow fever-carrying mosquitoes and non-human primate species richness). This approach is used to calculate epidemiological parameter values for a set of regions with known values of environmental covariates from a single set of coefficients.

For more information and background on the use of environmental covariates and the selection of covariates for use, see [The global burden of yellow fever (Gaythorpe et al, 2021, eLife)](https://doi.org/10.7554/eLife.64670).

Environmental data is stored as a data frame with each line containing covariate values for one region, with the headings giving the variable names. In this imaginary example, the environmental covariates are simply named Var1-5.

```{r,eval=FALSE}
library(YellowFeverDynamics)
enviro_data <- read.csv(file = paste(path.package("YellowFeverDynamics"), "/exdata/enviro_data_example.csv", sep = ""), header = TRUE)
head(enviro_data, 5)
```

Logarithmic values of the coefficients of the environmental covariates are stored as a data frame with the headings in the format FOI_(variable name) and R0_(variable name). In the example below, only one set of coefficient values are given, but multiple sets of coefficient values (e.g. a distribution of value estimates) can be used to produce a distribution of epidemiological parameter values, as shown below.

```{r,eval=FALSE}
enviro_coeffs <- read.csv(file = paste(path.package("YellowFeverDynamics"), "/exdata/enviro_coeffs_example.csv", sep = ""), header = TRUE)
enviro_coeffs
```

Covariate and coefficient values can be combined to produce values of the spillover force of infection and basic reproduction number using the param_calc_enviro() function.

```{r,eval=FALSE}
n_regions <- nrow(enviro_data)
epi_data <- data.frame(region = enviro_data$region, FOI = rep(NA, n_regions), R0 = rep(NA, n_regions))
for(n_region in 1:n_regions){
 FOI_R0_values <- param_calc_enviro(param = as.numeric(enviro_coeffs[1, ]), enviro_data = enviro_data[n_region, ])
 epi_data$FOI[n_region] <- FOI_R0_values$FOI
 epi_data$R0[n_region] <- FOI_R0_values$R0
}
head(epi_data, 5)
```

Spillover FOI and R0 data obtained in this way can be plotted on a map of the regions under consideration using the create_map function(), as can other regional data such as the values of the environmental covariates. Regional boundary data is loaded from one or more shapefiles of the format used by [GADM](https://gadm.org/download_country_v3.html) using the map_shapes_load() function

```{r,eval=FALSE}
regions=epi_data$region
shape_data <- map_shapes_load(regions, shapefiles=c(paste(path.package("YellowFeverDynamics"), "/exdata/shapefile_example.shp", sep = "")), region_label_type="GID_1")
colour_scheme=readRDS(file=paste(path.package("YellowFeverDynamics"), "exdata/colour_scheme_example.Rds", sep="/"))
colour_scale=colour_scheme$colour_scale
null1 <- create_map(shape_data, epi_data$FOI, scale=pretty(epi_data$FOI, 8), colour_scale, pixels_max=720, text_size=1, map_title="", legend_title="FOI", legend_position="topleft", legend_format="e", legend_dp=1, output_file=NULL)
null2 <- create_map(shape_data, epi_data$R0, scale=pretty(epi_data$R0, 8), colour_scale, pixels_max=720, text_size=1, map_title="", legend_title="R0", legend_position="topleft", legend_format="f", legend_dp=1, output_file=NULL)
```

To generate a distribution of spillover FOI and R0 values for each region, a distribution of environmental coefficient values can be used:

```{r,eval=FALSE}
enviro_coeffs2 <- read.csv(file = paste(path.package("YellowFeverDynamics"), "/exdata/enviro_coeffs_example2.csv", sep = ""), header = TRUE)
head(enviro_coeffs2, 5)
epi_data <- data.frame(region = enviro_data$region, FOI = rep(NA, n_regions), R0 = rep(NA, n_regions))
for(n_region in 1:n_regions){
 FOI_R0_values <- param_calc_enviro(param = as.numeric(enviro_coeffs2[1, ]), enviro_data = enviro_data[n_region, ])
 epi_data$FOI[n_region] <- FOI_R0_values$FOI
 epi_data$R0[n_region] <- FOI_R0_values$R0
}
```
