---
title: "Worked Example 1 - Running Time Delay Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example 1 - Running Time Delay Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This worked example demonstrates how to set up and run the time delay version of the standard yellow fever dynamic model presented in the [YEP package](https://mrc-ide.github.io/YEP/index.html). The time delay version differs from the standard version (which is described in [this worked example](https://mrc-ide.github.io/YEP/articles/BWorkedExampleAModelRun.html)) in using a fixed time delay to govern the movement of infected people from group E (exposed) to group I (infectious) and from group I to group R (recovered). In the standard model, this movement is governed by a rate given by the inverse of the time delay values (the sum of [TBA] in the case of E-I and the infectious period in the case of I-R). The [TBA] are all set to 5 days within the package code. [TBA - how to alter them]

Using a fixed delay period increases the complexity of the model (due to the properties of the architecture created using the [odin.dust package](https://mrc-ide.github.io/odin.dust/index.html)) but is potentially more accurate when modelling yellow fever transmission over short timescales.

The inputs for the Model_Run_Delay() function, which runs this version of the model, are the same as for the Model_Run function in the YEP package, except that there is no variable output type - the function always outputs detailed SEIRV data for every time point and age group (equivalent to setting output_type="full" in Model_Run()).

```{r,eval=FALSE}
library(YellowFeverDynamics)
input_data <- readRDS(file = paste(path.package("YellowFeverDynamics"), 
                                   "/exdata/input_data_example.Rds", sep = ""))
vacc_data <- input_data$vacc_data[1, , ]
pop_data <- input_data$pop_data[1, , ]  
# Force of infection due to spillover from sylvatic reservoir
FOI_spillover <- 1.0e-8             
# Basic reproduction number for urban spread of infection
R0 <- 1.5             
# Year(s) for which data is to be output
years_data <- c(2000)
# SEIRV data from end of a previous run to use as input (if mode_start = 2; not 
# used here and hence set to NULL)
start_SEIRV <- NULL           
# First year in population/vaccination data (here taken from the input data set; 
# if the population and vaccination data is set manually, this similarly needs 
# to be set manually)
year0 <- input_data$years_labels[1]
# Flag indicating how to set initial population immunity level in addition to
# vaccination
# If mode_start = 0, only vaccinated individuals are immune (R = 0)
# If mode_start = 1, shift some non-vaccinated individuals into recovered to 
# give herd immunity (stratified by age)
# If mode_start = 2, use SEIRV input in list from previous run(s), set using
# start_SEIRV variable
mode_start <- 1  
# Proportional vaccine efficacy (from 0 to 1); probability of vaccination 
# successfully providing immunity (i.e. moving vaccinated individual to group V)
vaccine_efficacy <- 1.0       
# Time increment in days to use in model (should be 1.0, 2.5 or 5.0 days)
time_inc <- 1.0            
# Number of particles to run
n_particles <- 10     
# Number of threads to run
n_threads <- 1   
# True/false flag indicating whether or not to run model in deterministic mode
# (so that binomial calculations give average instead of randomized output)
deterministic <- FALSE
```

We run Model_Run_Delay() as shown:

```{r,eval=FALSE}
test1 <- Model_Run_Delay(FOI_spillover, R0, vacc_data, pop_data, years_data,
                   start_SEIRV, year0, mode_start, vaccine_efficacy, 
                   time_inc, n_particles, n_threads, deterministic)
```

As with YEP::Model_Run(), we can plot the output:

```{r,eval=FALSE}
library(YEPaux)
plot <- plot_model_output(test1)
print(plot)
```
