---
title: "Worked Example 3 - Running Split Infection Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example 3 - Running Split Infection Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This worked example demonstrates how to set up and run the split infection version of the standard yellow fever dynamic model presented in the [YEP package](https://mrc-ide.github.io/YEP/index.html). The split infection version differs from the standard version (which is described in [this worked example](https://mrc-ide.github.io/YEP/articles/BWorkedExampleAModelRun.html)) in dividing infections into sylvatic and urban.

Sylvatic transmission refers to infection of humans via bites from mosquitoes infected with yellow fever via biting infected non-human primates (NHPs). Urban transmission refers to infection of humans via bites from mosquitoes infected with yellow fever via biting other humans.

The inputs for the Model_Run_Split() function, which runs this version of the model, are the same as for the Model_Run function in the YEP package, except that there is no variable output type - the function always outputs detailed SEIRV data for every time point and age group (equivalent to setting output_type="full" in Model_Run()).

```{r,eval=FALSE}
library(YellowFeverDynamics)
input_data <- readRDS(file = paste(path.package("YellowFeverDynamics"), 
                                   "/exdata/input_data_example.Rds", sep = ""))
vacc_data <- input_data$vacc_data[1, , ]
pop_data <- input_data$pop_data[1, , ]  
# Force of infection due to spillover from sylvatic reservoir
FOI_spillover <- 1.0e-8             
# Basic reproduction number for urban spread of infection
R0 <- 1.5             
# Year(s) for which data is to be output
years_data <- c(2000)
# First year in population/vaccination data (here taken from the input data set; 
# if the population and vaccination data is set manually, this similarly needs 
# to be set manually)
year0 <- input_data$years_labels[1]
# Proportional vaccine efficacy (from 0 to 1); probability of vaccination 
# successfully providing immunity (i.e. moving vaccinated individual to group V)
vaccine_efficacy <- 1.0      
# Time increment in days to use in model (should be 1.0, 2.5 or 5.0 days)
time_inc <- 1.0     
# Type of data to output
#'   "full" = SEIRVC + FOI for all steps and ages
#'   "case" = annual total new infections (C) summed across all ages
#'   "sero" = annual SEIRV
#'   "case+sero" = annual SEIRVC, cases summed across all ages
#'   "case_alt" = annual total new infections not combined by age
#'   "case_alt2" = total new infections combined by age for all steps
mode_start <- 1           
# SEIRV data from end of a previous run to use as input (if mode_start = 2; not 
# used here and hence set to NULL)
start_SEIRV <- NULL   
# Type of time variation of FOI_spillover and R0; here set to 0 for constant values
mode_time <- 0
# Number of particles to run
n_particles <- 10     
# Number of threads to run
n_threads <- 1   
# True/false flag indicating whether or not to run model in deterministic mode
# (so that binomial calculations give average instead of randomized output)
deterministic <- FALSE
```

We run Model_Run_Split() as shown:

```{r,eval=FALSE}
test1 <- Model_Run_Split(FOI_spillover, R0, vacc_data, pop_data, years_data, year0,
                         vaccine_efficacy, time_inc, mode_start, start_SEIRV, mode_time, 
                         n_particles, n_threads, deterministic)
```

We can plot the output and compare the number of sylvatic and urban infections:

```{r,eval=FALSE}
dates=test1$day-test1$day[1]+1
C_sylvatic = C_urban = rep(NA, length(dates)) # Cumulative sylvatic and urban infections
C_sylvatic[1] = sum(test1$C_sylvatic[,,1])/n_particles
C_urban[1] = sum(test1$C_urban[,,1])/n_particles
for(i in 2:length(dates)){
  C_sylvatic[i] = C_sylvatic[i-1] + sum(test1$C_sylvatic[,,i])/n_particles
  C_urban[i] = C_urban[i-1] + sum(test1$C_urban[,,i])/n_particles
}
matplot(x = dates, y = C_sylvatic, ylim = c(0, max(C_sylvatic,C_urban)), type = "l", col = 1,
        xlab = "Date", ylab = "Cumulative infections")
matplot(x = dates, y = C_urban, type = "l", col = 2, add = TRUE)
legend("topleft",legend = c("Sylvatic","Urban"), lty = c(1,1), col = c(1,2))
```
