---
title: "Worked Example 1 - Single Model Run"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example 1 - Single Model Run}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This worked example demonstrates how to set up and carry out a single run of the model for a single region. 

The model of yellow fever used in YellowFeverDynamics is an age-stratified, stochastic SEIRV model in which the number of people in each age group in groups S (susceptible), E (exposed), I (infectious), R (recovered) and V (vaccinated) is computed at each time point, with changes in population also incorporated. The number of new infections (individuals moving from group S to group E) at each time point is calculated stochastically using the total force of infection (FOI) as a binomial probability. Full details of the model are currently being prepared for publication; please contact [Keith Fraser](https://www.imperial.ac.uk/people/keith.fraser) for a detailed description.

Population and vaccination data is stored in a list object as described in [Guide 1 - Inputs](https://mrc-ide.github.io/YellowFeverDynamics/articles/CGuideAInputs.html). Here we load an example data set for multiple regions and extract the vaccination and population data for the first included region:

```{r}
library(YellowFeverDynamics)
input_data <- readRDS(file = paste(path.package("YellowFeverDynamics"), "/exdata/input_data_example.Rds", sep = ""))
vacc_data <- input_data$vacc_data[1, , ]
pop_data <- input_data$pop_data[1, , ]  
```

We set the other input parameters to the functions which run the model as follows. Here we set the spillover FOI and basic reproduction number directly; they may alternatively be calculated from environmental covariates as described in [Guide 2 - Calculating Parameters From Environmental Data](https://mrc-ide.github.io/YellowFeverDynamics/articles/CGuideBParametersEnvironment.html). The mode_start variable is used to set the initial conditions in one of 3 ways. The n_particles variable is used to run the model multiple times for the same inputs; the results can be compared and/or combined to test and/or average out the effects of the stochasticity of the model. The n_threads variable is used to set the number of particles run in parallel (subject to the limitations of the computer being used).

```{r}
FOI_spillover <- 1.0e-8             # Force of infection due to spillover from sylvatic reservoir
R0 <- 1.5                           # Basic reproduction number for urban spread of infection
year0 <- input_data$years_labels[1] # First year in population/vaccination data (here taken from the input data set; if the population and vaccination data is set manually, this similarly needs to be set manually)
mode_start <- 1                     # Flag indicating how to set initial population immunity level in addition to
                                    # vaccination
                                    # If mode_start = 0, only vaccinated individuals are immune (R = 0)
                                    # If mode_start = 1, shift some non-vaccinated individuals into recovered to give
                                    # herd immunity (R = 1 - 1/R0)
                                    # If mode_start = 2, use SEIRV input in list from previous run(s), set using
                                    # start_SEIRV variable
n_particles <- 10                   # Number of particles to run (maximum allowed value = 20)
n_threads <- 1                      # Number of threads to run (equal to or less than n_particles)
year_end <- 2001                    # Year at which to stop generating data (year after last year where data is desired)
year_data_begin <- 1999             # Year to begin saving data
vaccine_efficacy <- 1.0             # Proportional vaccine efficacy (from 0 to 1)
start_SEIRV <- NULL                 # SEIRV data from end of a previous run to use as input (if mode_start = 2; not used here and hence set to NULL)
dt <- 1.0                           # Time increment in days to use in model (should be either 1.0 or 5.0 days)
```

We use the Full_Model_Run() function to run the model for the inputs above:

```{r}
test1 <- Full_Model_Run(FOI_spillover, R0, vacc_data, pop_data, year0, mode_start, n_particles, n_threads, year_end,
                        year_data_begin, vaccine_efficacy, start_SEIRV, dt)
```

The output of Full_Model_Run() includes, for each particle, date information, the total force of infection FOI_total at each data point, the numbers of people in groups S, E, I, R and V at each time point organized by age group, and the number of newly infectious people (C) at each time point organized by age group. We use the plot_model function to display values of S, R and V (added together across all age groups) over time, with error bars to show the variation between particles caused by stochasticity.

```{r}
plot<-plot_model_output(test1)
print(plot)
```

For some applications, we can use the Basic_Model_Run() function instead. This is largely identical to Full_Model_Run() but does not output the total force of infection at each time point or the number of new infections at each time point. This reduces the amount of data output by each run of the function, reducing computational requirements. For example, we use Basic_Model_Run when the intention is only to calculate seroprevalence, which is derived (as described in [Guide 4 - Processing Model Results](https://mrc-ide.github.io/YellowFeverDynamics/articles/BGuideDModelResultsProcess.html)) from SEIRV values, with the rate of newly infectious individuals being irrelevant.

```{r,eval = FALSE}
test2 <- Basic_Model_Run(FOI_spillover, R0, vacc_data, pop_data, year0, mode_start, n_particles, n_threads, year_end,
                         year_data_begin, vaccine_efficacy, start_SEIRV, dt)
```

For applications where only the number of new infectious people at each time point is desired, we use the case_data_generate() function instead. This functions similarly to Basic_Model_Run() and Full_Model_Run() but only outputs data on the total number (across all age groups) of newly infectious people at each time point. It is also designed to run large numbers of repetitions, automatically sequentially running smaller groups of particles and threads to avoid causing errors or freezing the computer while running (Basic_Model_Run() and Full_Model_Run() are limited to 20 particles/threads for this reason). The ability to run a large number of repetitions is valuable when modelling annual case rates and/or outbreak risk/occurrence (see [Guide 3 - Processing Model Results](https://mrc-ide.github.io/YellowFeverDynamics/articles/CGuideCModelResultsProcess.html)) in situations where the overall force of infection is low and therefore new infections are rare. Here we run case_data_generate() for 1000 repetitions with a low spillover FOI and no human-to-human transmission, which gives a very low daily infection rate.

```{r,eval = FALSE}
n_reps <- 1000
FOI_spillover <- 1.0e-10
R0 <- 0.0
test3 <- case_data_generate(FOI_spillover, R0, vacc_data, pop_data, year0, mode_start, n_reps, year_end,
                            year_data_begin, vaccine_efficacy, start_SEIRV, dt)
```
