---
title: "Worked Example 2 - Running Reactive Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example 2 - Running Reactive Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This worked example demonstrates how to set up and run the reactive version of the standard yellow fever dynamic model presented in the [YEP package](https://mrc-ide.github.io/YEP/index.html). The reactive version differs from the standard version (which is described in [this worked example](https://mrc-ide.github.io/YEP/articles/BWorkedExampleAModelRun.html)) in including provision for an emergency response to reported cases and/or high burden levels. This version of the model can be used, for example, to evaluate the optimum emergency or preventive measures to prevent large outbreaks.

The inputs for the Model_Run_Delay_Reactive() function, which runs this version of the model, are the same as for the Model_Run_Delay function detailed in [Worked Example 1](https://mrc-ide.github.io/YellowFeverDynamics/articles/BWorkedExampleAModelRunDelay.html), with several additional inputs governing the detection of a significant yellow fever outbreak and the resulting response.

```{r,eval=FALSE}
library(YellowFeverDynamics)
input_data <- readRDS(file = paste(path.package("YellowFeverDynamics"), 
                                   "/exdata/input_data_example.Rds", sep = ""))
vacc_data <- input_data$vacc_data[1, , ]*0
pop_data <- input_data$pop_data[1, , ]  
# Force of infection due to spillover from sylvatic reservoir
FOI_spillover <- 1.0e-8 
# Basic reproduction number for urban spread of infection
R0 <- 1.0             
# Year(s) for which data is to be output
years_data <- c(1940:1945)        
# First year in population/vaccination data (here taken from the input data set; 
# if the population and vaccination data is set manually, this similarly needs 
# to be set manually)
year0 <- input_data$years_labels[1]
# Proportional vaccine efficacy (from 0 to 1); probability of vaccination 
# successfully providing immunity (i.e. moving vaccinated individual to group V)
vaccine_efficacy <- 1.0      
# Time increment in days to use in model (should be 1.0, 2.5 or 5.0 days)
time_inc <- 1.0     
mode_start <- 1           
# SEIRV data from end of a previous run to use as input (if mode_start = 2; not 
# used here and hence set to NULL)
start_SEIRV <- NULL   
# Type of time variation of FOI_spillover and R0; here set to 0 for constant values
mode_time <- 0
# Number of particles to run
n_particles <- 1   
# Number of threads to run
n_threads <- 1   
# True/false flag indicating whether or not to run model in deterministic mode
# (so that binomial calculations give average instead of randomized output)
deterministic <- FALSE
# Delay time between an event which triggers the declaration of an outbreak and
# the response (emergency vaccination campaign)
response_delay <- 56
# Probabilities of an infection being reported as a case before emergency conditions
# triggered (1st value) or after emergency conditions triggered (2nd value)
p_rep <- c(1,1)
# Threshold total no. reported cases to trigger emergency conditions
case_threshold <- 1.0
# Threshold current infectious fraction to trigger emergency conditions
cluster_threshold <- Inf
# Vaccination coverage by age group of emergency vaccination campaign
vacc_cov_cam <- c(0,rep(0.8,60),rep(0,40))
# Duration in days of emergency vaccination campaign
t_cam <- 60
```

We run Model_Run_Reactive() as shown:

```{r,eval=FALSE}
set.seed(1)
test1 <- Model_Run_Delay_Reactive(FOI_spillover, R0, vacc_data, pop_data, years_data, year0,
                                  vaccine_efficacy, time_inc, mode_start, start_SEIRV, mode_time, 
                                  n_particles, n_threads, deterministic, response_delay, p_rep,
                                  case_threshold, cluster_threshold, vacc_cov_cam, t_cam)
```

We can plot the output, including the application of the emergency vaccination campaign when the first case is reported:

```{r,eval=FALSE}
matplot(x=test1$day,y=t(test1$C_rep_total), type="l",xlab="Day",ylab="Reported cases (cumulative)")
```
