---
title: "Vignette 4 - Running Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette 4 - Running Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to set up and carry out a single run of the model for a single region. 

The model of yellow fever used in YellowFeverDynamics is an age-stratified, stochastic SEIRV model in which the number of people in each age group in groups S (susceptible), E (exposed), I (infectious), R (recovered) and V (vaccinated) is computed at each time point, with changes in population also incorporated. The number of new infections (individuals moving from group S to group E) at each time point is calculated stochastically using the total force of infection (FOI) as a binomial probability. Full details of the model are currently being prepared for publication; please contact [Keith Fraser](keith.fraser@imperial.ac.uk) for a detailed description.

First load the package:

```{r}
library(YellowFeverDynamics)
```

Population and vaccination data is stored in a list object as described in Vignette 2. Here an example data set for multiple regions is loaded and the vaccination and population data for the first included region is extracted:

```{r}
input_data <- readRDS(file = paste(path.package("YellowFeverDynamics"),"/exdata/input_data_example.Rds",sep = ""))
vacc_data <- input_data$vacc_data[1,,]
pop_data <- input_data$pop_data[1,,]  
```

The other input parameters to the functions which run the model are as follows. Here the spillover FOI and basic reproduction number are set directly; they may alternatively be calculated from environmental covariates as described in Vignette 3.

```{r}
FOI_spillover <- 1.0e-8            # Force of infection due to spillover from sylvatic reservoir
R0 <- 1.5                          # Reproduction number for urban spread of infection
year0 <- 1940                      # First year in population/vaccination data
mode_start <- 1                    # Flag indicating how to set initial population immunity level in addition to vaccination
                                   # If mode_start = 0, only vaccinated individuals
                                   # If mode_start = 1, shift some non-vaccinated individuals into recovered to give herd immunity
                                   # If mode_start = 2, use SEIRVC input in list from previous run(s)
n_particles <- 10                  # Number of particles to run (maximum allowed value = 20)
n_threads <- 1                     # Number of threads to run (equal to or less than n_particles)
year_end <- 2001                   # Year at which to stop generating data (year after last year where data is desired)
year_data_begin <- 1999            # Year to begin saving data
vaccine_efficacy <- 1.0            # Proportional vaccine efficacy
start_SEIRV <- NULL                # SEIRV data from end of a previous run to use as input (if mode_start = 2)
dt <- 1.0                          # Time increment in days to use in model (should be either 1.0 or 5.0 days
```

The Full_Model_Run function is normally used to run the model for the inputs above:

```{r}
test1 <- Full_Model_Run(FOI_spillover,R0,vacc_data,pop_data,year0,mode_start,n_particles,n_threads,year_end,
                        year_data_begin,vaccine_efficacy,start_SEIRV,dt)
```

The output of Full_Model_Run includes, for each repetition, date information, the total force of infection FOI_total at each data point, the numbers of people in groups S, E, I, R and V at each time point and the number of newly infectious people (C) at each time point organized by age group.

```{r}
plot<-plot_model_output(test1)
print(plot)
```

For some applications, the Basic_Model_Run function can be used instead. This is identical to Full_Model_Run but does not output the total force of infection at each time point or the number of new infections at each time point.

```{r,eval = FALSE}
test2 <- Basic_Model_Run(FOI_spillover,R0,vacc_data,pop_data,year0,mode_start,n_particles,n_threads,year_end,
                     year_data_begin,vaccine_efficacy,start_SEIRV,dt)
```

For applications where only the number of daily infections is desired, the case_data_generate function can be used instead. This functions similarly to Basic_Model_Run and Full_Model_Run but only outputs infection data. It is also designed to run large numbers of repetitions, automatically sequentially running smaller groups of particles and threads to avoid causing errors or freezing the computer while running. The ability to run a large number of repetitions is valuable when modelling outbreak occurrence in situations where the overall force of infection is low. Hence in the example below, case_data_generate is run for 1000 repetitions with a low spillover FOI and no human-to-human transmission, resulting in a very low daily infection rate.

```{r,eval = FALSE}
n_reps <- 1000
FOI_spillover <- 1.0e-10
R0 <- 0.0
test3 <- case_data_generate(FOI_spillover,R0,vacc_data,pop_data,year0,mode_start,n_reps,year_end,year_data_begin,
                            vaccine_efficacy,start_SEIRV,dt)
```
