---
title: "Vignette 5 - Processing Model Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette 5 - Processing Model Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how the output of the dynamic yellow fever model functions (see vignette 4) can be processed to produce epidemiological data similar to that produced by real-world surveillance. There are three types of data processed by the package:

-Serological survey data, the proportion of people tested in a chosen age range who test positive for yellow fever antibodies (indicating that they have either recovered from the disease or been vaccinated, with individuals known to have been vaccinated normally excluded from surveys - see below for more on this) [TBA - Link to example publications].
-Annual reported cases and deaths, representing the number of laboratory-confirmed yellow fever cases in a region in a given year and the number which resulted in death. These usually represent a fraction of the actual number of annual yellow fever infections and deaths in a region (in particular, it is usually assumed that infections causing mild or no symptoms are never reported as non-fatal cases cases).
-Annual outbreak occurrence, binary data indicating whether one or more outbreaks (usually defined as one or more lab-confirmed cases) are reported in a region in a given year.

A set of epidemiological data for multiple regions can be constructed using additional functions described in vignette 6; this combines running the model as described in vignette 4 and calculating epidemiological output data as described in this vignette.

Data is generated as described in vignette 4, in this case using the Full_Model_Run() function in order to be able to calculate all three epidemiological data types.

```{r}
library(YellowFeverDynamics)
input_data <- readRDS(file=paste(path.package("YellowFeverDynamics"),"/exdata/input_data_example.Rds",sep=""))
vacc_data <- input_data$vacc_data[1,,]
pop_data <- input_data$pop_data[1,,]  
FOI_spillover <- 1.0e-8             
R0 <- 1.5                           
year0 <- input_data$years_labels[1] 
mode_start <- 1                     
n_particles <- 10                   
n_threads <- 1                      
year_end <- 2001                    
year_data_begin <- 1990
vaccine_efficacy <- 1.0             
start_SEIRV <- NULL
dt <- 1.0
model_data <- Full_Model_Run(FOI_spillover,R0,vacc_data,pop_data,year0,mode_start,n_particles,n_threads,year_end,year_data_begin,vaccine_efficacy,start_SEIRV,dt)
```

The seroprevalence in a given age range in one or more years for one particle may be calculated from model data using the sero_calculate() function as shown below (in this case, for 0-15 year olds from 1990-2000). Seroprevalence for a given year calculated in this way is a year-round average. The variable vc_factor is used when creating modelled seroprevalence data to compare with real-world surveys, to account for the fact that in real-world seroprevalence surveys, while normally only unvaccinated individuals are tested, some surveyed individuals' vaccination status may be unknown. If vc_factor = 0 (i.e. all individuals tested are confirmed unvaccinated), seroprevalence is calculated only for the unvaccinated population. If vc_factor > 0, seroprevalence is calculated for an appropriate mixture of the unvaccinated population (where seroprevalence = R/(S+E+I+R)) and the overall population (where seroprevalence = (R+V)/(S+E+I+R+V)).

```{r}
age_min <- 0      # Minimum age in age range to survey
age_max <- 15     # Maximum age in age range to survey
years <- c(1990:2000)  # Year(s) for which to calculate seroprevalence
vc_factor <- 0    # Fraction of the population (from 0 to 1) for whom vaccination status is unknown
n_particle <- 1   # Particle to select from model_data
seroprevalence <- sero_calculate(age_min,age_max,years,vc_factor,model_data,n_particle)
matplot(x=years,y=seroprevalence,type="b",pch=1,xlab="Year",ylab="Seroprevalence")
```

Annual reported severe and fatal case data is generated from C, the output data on the number of new infectious individuals at each time point. The actual number of severe and fatal cases is calculated stochastically using the variables p_severe_inf (the probability of an infection leading to severe symptoms) and p_death_severe_inf (the probability that a severe infection with lead to death). The values set for these below (0.12 and 0.47, taken from [TBA - link to source for values]) are incorporated into the package as global variables. The number of reported severe and fatal cases is then calculated stochastically using the variables p_rep_severe (probability of a severe case being reported in public health figures, usually meaning lab-confirmed cases) and p_rep_death (probability of a fatal case being reported). The total number of reported cases (rep_cases in the output data frame annual_data below) is the sum of reported severe cases and reported fatal cases (so it will always be equal to or greater than the number of reported deaths).

```{r}
p_severe_inf=0.12
p_death_severe_inf=0.47
p_rep_severe=0.05
p_rep_death=0.1
n_years=length(years)
annual_data=list(rep_cases=rep(NA,n_years),rep_deaths=rep(NA,n_years))
for(n_year in 1:n_years){
  infs=sum(model_data$C[,n_particle,model_data$year[1,]==years[n_year]])
  severe_infs=rbinom(1,floor(infs),p_severe_inf)
  deaths=rbinom(1,severe_infs,p_death_severe_inf)
  annual_data$rep_deaths[n_year]=rbinom(1,deaths,p_rep_death)
  annual_data$rep_cases[n_year]=annual_data$rep_deaths[n_year]+rbinom(1,severe_infs-deaths,p_rep_severe)
}
matplot(x=years,y=annual_data$rep_cases,type="b",pch=1,xlab="Year",ylab="Individuals")
matplot(x=years,y=annual_data$rep_deaths,type="b",pch=1,col=2,add=TRUE)
legend("topright",legend=c("Reported cases","Reported deaths"),lty=c(1,1),pch=c(1,1),col=c(1,2))
```

Outbreak occurrence data can be calculated from annual reported case data; if the case count for a given year is greater than zero, it is flagged as a year in which one or more outbreaks will be reported and the outbreak occurrence value is set to 1; otherwise, it is 0.

```{r}
outbreak_occurrence=rep(0,n_years)
for(n_year in 1:n_years){
  if(annual_data$rep_cases[n_year]>0){outbreak_occurrence[n_year]=1}
}
barplot(height=outbreak_occurrence,names.arg=years)
```

When the model is run for multiple particles, outbreak risk can be calculated for each year from the proportion of particles in which the year is flagged as an outbreak year.

```{r}
n_reps <- 1000
FOI_spillover <- 1.0e-8
R0 <- 0.0
model_data2 <- case_data_generate(FOI_spillover,R0,vacc_data,pop_data,year0,mode_start,n_reps,year_end,year_data_begin,vaccine_efficacy,start_SEIRV,dt)

outbreak_risk=rep(0,n_years)
dp=1/n_reps
for(n_rep in 1:n_reps){
  for(n_year in 1:n_years){
    infs=sum(model_data2$C[n_particle,model_data2$year==years[n_year]])
    severe_infs=rbinom(1,floor(infs),p_severe_inf)
    deaths=rbinom(1,severe_infs,p_death_severe_inf)
    rep_deaths=rbinom(1,deaths,p_rep_death)
    rep_cases=rep_deaths+rbinom(1,severe_infs-deaths,p_rep_severe)
    if(rep_cases>0){outbreak_risk[n_year]=outbreak_risk[n_year]+dp}
  }
}
matplot(x=years,y=outbreak_risk,type="b",pch=1,xlab="Year",ylab="Outbreak risk")
```
