---
title: "ModelRun"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ModelRun}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to set up and carry out a single run of the model for a single region. 

First load the package:

```{r}
library(YellowFeverDynamics)
```

[TBA]

```{r,eval=FALSE}
input_data_all <- readRDS(file=paste(path.package("YellowFeverDynamics"),"/exdata/input_data_example.Rds",sep=""))
regions=input_data_all$region_labels[1]
check=input_data_all$region_labels %in% regions
input_data=list(region_labels=input_data_all$region_labels[check],years_labels=input_data_all$years_labels,
                age_labels=input_data_all$age_labels,vacc_data=input_data_all$vacc_data[check,,],
                pop_data=input_data_all$pop_data[check,,])
```

[TBA]

```{r,eval=FALSE}
FOI_spillover=1.0e-6            #
R0=1.0                          #
vacc_data=input_data$vacc_data  #
pop_data=input_data$pop_data    #
year0=1940                      #
mode_start=1                    #
n_particles=1                   #
n_threads=1                     #
year_end=2000                   #
year_data_begin=1999            #
vaccine_efficacy=1.0            #
start_SEIRVC=NULL               #
dt=1.0                          #
```

The Full_Model_Run function is normally used to run the model for the inputs above:

```{r,eval=FALSE}
test1<-Full_Model_Run(FOI_spillover,R0,vacc_data,pop_data,year0,mode_start,n_particles,n_threads,year_end,
                     year_data_begin,vaccine_efficacy,start_SEIRVC,dt)
```

For some applications, the Basic_Model_Run function can be used instead. This is identical to Full_Model_Run but does not output the total FOI at each time point or the number of new infections at each time point.

```{r,eval=FALSE}
test2<-Basic_Model_Run(FOI_spillover,R0,vacc_data,pop_data,year0,mode_start,n_particles,n_threads,year_end,
                     year_data_begin,vaccine_efficacy,start_SEIRVC,dt)
```
